### 2019-01-15 sections/conclusion.tex

\section{Conclusion and Future Work}

In this paper we have shown how to mechanise an algorithmic subtyping relation
for higher-order polymorphism, together with its proofs of soundness,
completeness and decidability, in the Abella proof assistant. 
In ongoing work we are extending our mechanisation with a bidirectional type
inference algorithm. The main difficulty there is communicating the
instantiations of existential variables from the subtyping algorithm to the
type inference. To make this possible we are exploring a continuation passing
style formulation, which generalises the worklist approach.
Another possible extension is to have the algorithm return an explicit witness
for the subtyping as part of type-directed elaboration into System F.