### 2019-02-11 sections/conclusion.tex
0211 TODOs

\section{Conclusion}

This is the conclusion.

\jimmy{Discuss the efficient algorithm and its equivalence}