### 2019-02-27 sections/Appendix.tex
create the appendix tex file

\section{Appendix}

\jimmy{Lemma 29 is false. TODO structure and more text}
The first conclusion of the lemma says that if $\Gm_L\longrightarrow \Delta_R$ and
$\Gm_L,\Gm_R\longrightarrow \Delta_L,\Delta_R$ then
$\Gm_L,\al,\Gm_R\longrightarrow \Delta_L,\al,\Delta_R$.
If we pick $\Gm_L = \nil, \Gm_R = \bt, \Delta_L = \bt, \Delta_R = \nil$, the conditions
$\nil\longrightarrow \bt$ and $\bt\longrightarrow \bt$ hold, but the conclusion
$\al,\bt \longrightarrow \bt,\al$ does not hold.