### 2019-03-01 sections/conclusion.tex
Outline of S7 Conclusion

\section{Conclusion}

This is the conclusion.

% \jimmy{Discuss the efficient algorithm and its equivalence}
% The algorithm is already efficient

Outline:
\begin{itemize}
    \item difficulties following DK's formalization
    \item a blend of ideas from various sources DK's ordered contexts
    \item list of judgements + some novel ideas
    \item Future work: 
\end{itemize}