### 2019-04-18 response.md
reply

Response for Submission 66, ICFP 2019
======

We are grateful for the helpful reviews and are happy
that you appreciate our work. In what follows we
respond to your main comments.

Review A
------
<!--
One minor point of criticism I have is that there appear to be some
unexplained differences between the presentation in the paper and the
Abella mechanization, which made it harder than necessary for me to
relate the two and to follow some of the proof sketches in the
paper. For example, the (binary) multi-step reduction relations on
algorithmic and declarative worklists seem to be represented by unary
predicates (judge and dc respectively) in the Abella script, which
look more like big-step than small-step reduction relations. This
change of representation may even impact the structure of some proofs
(e.g. the induction hypothesis used in the proofs of soundness and
completeness). I know very little about Abella, so this choice of
representation may well be more idiomatic than the one presented in
the paper. Still, I would have liked to see a short discussion of the
pros/cons of the two representations and why different representations
are used in the paper and the proof script. More extensive
documentation of the proof script (e.g. more comments) would also be
helpful in navigating the mechanization.
-->

> The difference between the presentation in the paper and the Abella mechanization.

In the paper, the main algorithmic rules are written in a simple small
step style with no premise on the reduction relation (-->) itself.
And the rules are not overlapping with each other.  These properties
indicates that a trivial translation to the big step style used in the
proofs preserves equivalence.

We believe the presentation of the algorithmic rules in the paper is
much clearer and restricted to the form $\Gamma$ --> $\Gamma$, while
the big step unary relation in the proof simplifies the induction
scheme.  More importantly, the difference is just a matter of
presentation, but the logic behind remains unchanged.
We will mention this difference in presentation clearly in the paper.

> The restriction of subtyping to simple  types is effectively an
> equivalence (and thus symmetric).  Is this correct?

Yes, this is correct. Note also that DK's algorithm also suggests a
similar symmetric unification.

> The proof scripts are hard to read, and some parts are inconsistent with the paper.

We apologize for the state of the proof scripts. Currently, we are refactoring them.
and will document the mapping between the paper and the mechanisation
in the final version of the paper.

Review B
------
<!--
On the negative part, I am not sure that one can, from the paper, infer general hints/guidelines about the formalization of the correctness/completeness/... of an inference algorithm for a complex typing discipline. This is mainly due to the fact that the authors concentrate their attention to one single inference algorithm.
-->
<!--
Tom:
- Apparently Reviewer C believes they will be able to try out our approach on new problems.
- I also think there is an obvious generalization to other bidirectional type systems.
- We can say that we will extend the discussion to point this out.
-->

> General hints/guidlines about the formalization of the properties of
> an inference algorithm for a complex typing descipline.

That is a good point. We believe that our approach can indeed be
adapted to other bidirectional type systems, and we will extend the
discussion in the paper to point this out.  We are also happy to see
that Reviewer C already looks forward to trying out our approach in
other settings.